Empirical study of land surface effects on local rainfall

### 2018-09-04 Manuscripts/chap3_paper_revision.Rnw
changes for paper

\begin{abstract}
(text here)
\end{abstract}

@@

Land use and land cover changes can lead to changes in the local climate. Both empirical and modelling studies have found cloud types and rainfall are correlated to large scale vegetation cover changes, such as deforestation in the Amazon and in the Sahel \citep{Chagnon2005,Pinto2009, Wang2009, Mei2010, kucharski_further_2013,pitman_scale_2016} and afforestation in south Israel \citep{Otterman1990, Ben-Gai1998}. Using airborne measurement in Western Australia, \citet{Junkermann2009} showed a significantly higher level of aerosols over an agricultural area compared to an adjacent natural vegetation. They suggested that a modification of aerosol concentrations due to deforestation could have contributed to a reduction of local rainfall, as more but smaller rain droplets were observed. \citet{Nair2011} reported from the Bunny Fence Experiment in Western Australia that local land use change altered the synoptic west coast trough dynamics and surface roughness, and this resulted in the observed rainfall decrease. Climate sensitivity to land cover change is also found in eastern Australia \citep{McAlpine2007}.

Rainfall over land is generally influenced by multiple factors. Locally, there are two main sources: moisture from advective atmospheric transport; and local evapotranspiration \citep{Eltahir1996,Bosilovich2006,Dirmeyer2009,Gimeno2010}. According to \citet{Trenberth1999}, the contribution of advective moisture partially depends on the availability of external moisture and atmospheric transport. On the longer time scale, such as monthly and annually, large scale atmospheric dynamics are affected by large scale climate drivers. Many studies have reported significant relationships between rainfall in large parts of Australia and the El Ni\~{n}o-Southern Oscillation (ENSO) \citep{Verdon2004,Risbey2009,Speer2011}. ENSO can be used to represent longer term cycles in rainfall data such as drought. On the other hand, local ET is determined by local land surface characteristics. Local land surface characteristics further influence local scale atmospheric dynamics and hence the amount of rainfall, including contribution from both sources. Therefore land surface plays an important role in local rainfall.

Conventionally, large scale climate drivers are used as predictors for seasonal rainfall forecasts. According to the Australian Bureau of Meteorology (BoM), probablistic seasonal outlooks are developed based on Australian rainfall and temperature as well as sea surface temperature records from the tropical Pacific and Indian Oceans \citep{BoM2012c}. 
%% comment Expand this section with section from methods (section 4.1)?
Different parts of the Australian continent can be more or less influenced by different climatic drivers \citep{Mla2008}. Using both SOI and PDO in a prediction model, \citet{Kamruzzaman2011} reported that PDO was rarely significant for rainfall stations in the MDB and on the southeast coast of Australia, while SOI was at least significant at the 1\% level. In addition to SOI, \citet{Speer2011} found that the observed rainfall decrease in the southeast of NSW was linked to an increasingly positive SAM in 1976 - 1998. In southwest WA, the influence of any oceanic indices is small \citep{Smith2012}.

Although climate drivers demonstrate some capability to predict Australian rainfall, there is still a large amount of unexplained variance. \citet{Westra2010} pointed out that models based on global sea surface temperature anomalies can only predict up to 14.7\% of precipitation variance. Recent studies suggest that land surface processes are important for predicting local rainfall \citep[e.g.][]{Ma2011,Zeng2012, pitman_scale_2016, saha_investigating_2016}. However, they are mostly based on modelling experiments and little evidence was reported from observations. \citet{Pitman2004} found a good match between observations and simulated rainfall changes in southwest Western Australia ,forced by land cover change. \citet{Timbal2006} were able to reproduce the rainfall decline in south west Australia by including land cover influence. Local land use change might not be a primary, but is likely to be a secondary cause of rainfall change \citep{Nicholls2006}. Therefore, land surface modification has, at least partially, contributed to local rainfall variability.

Overall the number of empirical studies related to changes to rainfall due to land cover change is limited. This is because there are some fundamental experimental difficulties in both space (where does evaporated water reappear as rainfall?) and in time (how much time does it take for land cover change effects to appear or disappear?). Therefore, the aim of this study is to use empirical evidence at regional scales to investigate the cause and effect relationship between land cover change and local rainfall using empirical evidence. More specifically, we hypothesize that a step change on the land cover on the surface will cause a step change in the rainfall. To demonstrate the approach and this effect we studied step changes in rainfall at two locations in Queensland and NSW/Victoria where there are possible step changes in land cover change due to land clearing and bushfires based on tree cover data. Statistical approaches were used to identify the step changes in rainfall, which were subsequently associated with land cover change through spatial comparison.

In this paper, after this section (the introduction), section 2 covers the case study areas and the observed land use changed. Section 3 describes the data used in the study in more detail. Section 4 details the statistical methods and the underlying assumptions related to the modelling approach, Section 5 gives the results, which are further discussed in section 6 and finally section 7 offers the conclusions.

@@

In Australia, significant tree cover change has mainly occurred in the north east of the continent and on the southeast coast, as well as in the southwest of Western Australia. According to the National Dynamic Land Cover Dataset (DLCD) \citep{Lymburner2010}, most of these areas have experienced decreasing EVI between 2000 - 2008. As an index for vegetation greenness, the decreasing values indicate lower biomass over time in the tree cover regions. The possible EVI reduction might be due to land clearing, bushfires or drought.

@@

Two regions were selected where significant tree cover change was present. The first region is located in south central Queensland to the north of the Murray Darling Basin (MDB) (site 1 in Figure~\ref{fig:sel_reg}). High rates of land clearing have been reported in this region during the early 2000s \citep{SLATS2004}. The second study region is located at the border of New South Wales and Victoria, and includes the Snowy Mountains ranges (site 2 in Figure~\ref{fig:sel_reg}). Severe bushfires occurred in this area and the surroundings in early 2003 (see Figure~\ref{fig:bushfire}). The 2003 bushfires were the largest and the worst in this area for the last 60 years \citep{Fire2011}. Two thirds of Kosciuszko national park was heavily burned and regrowth was reported to be slow due to drought and cold conditions \citep{ABC2003} and the type of species in this region. In both study regions, significant tree cover loss has happened within the last decade, either permanently or temporarily. It makes the selected areas suitable for this land cover change study. 
%\footnote{Bushfire history, \url{http://www.dse.vic.gov.au/fire-and-other-emergencies/major-bushfires-in-victoria}, viewed on 12/12/2011}
%\footnote{Kosciuszko slow to recover from bushfires, \url{http://www.abc.net.au/news/2003-10-30/kosciuszko-slow-to-recover-from-bushfires/1501722}, viewed on 06/06/2012}

@@

<<mapping,eval=F, echo = F, results = hide>>=

library(plotKML)
library(raster)
library(rgeos)
library(maptools)
library(fields)
library(RColorBrewer)

## define the boundary of sub-regions
  e1 <- extent(c(1324304.091,1723901.773,-222556.668,80451.265))
  e2 <- extent(c(1346395.141,1534425.527,-1169402.246,-607644.171))

  e11 <- extent(c(145.425, 149.025, -28.475, -25.175))
  e22 <- extent(c(146.675, 149.325, -37.075, -34.675))

  e111 <- extent(c(145.825, 148.625, -28.075, -25.575)) #as(tc_r1,"SpatialPolygons")
  e222 <- extent(c(147.075, 148.925, -36.675, -35.075)) #as(tc_r2,"SpatialPolygons")

## the selected regions for mapping

  aus_reg <- readShapePoly("Data\\Geo\\aus_state2011\\simplify_state11_uproj.shp",proj4string=CRS("+proj=lcc +lat_1=-18 +lat_2=-36 +lat_0=-27 +lon_0=132 +x_0=0 +y_0=0"))
  aus_reg2 <- gSimplify(aus_reg,0.01,topologyPreserve=T)

  mdb_reg <- readShapePoly("Data\\Geo\\MDB\\mdb.shp",proj4string=CRS("+proj=lcc +lat_1=-18 +lat_2=-36 +lat_0=-27 +lon_0=132 +x_0=0 +y_0=0"))
mdb_reg2 <- gSimplify(mdb_reg,0.01,topologyPreserve=TRUE)

  remove(list=c("aus_reg","mdb_reg"))

@@

fire_0301 <- raster("Data\\Other\\bushfires\\Burnt_Area\\tmpA2003001.burndate.tif")
fire_0301nsw <- crop(fire_0301,e2)
fire_loc <- crop(reproject(fire_0301nsw,CRS=CRS("+proj=longlat +lat_1=-18 +lat_2=-36 +lat_0=-27 +lon_0=132 +x_0=0 +y_0=0 +ellps=WGS84")),e222)
remove(list=c("fire_0301","fire_0301nsw"))

  setEPS()
  postscript("bushfire_nswvic.eps")
#  jpg("bushfire_nswvic.jpg")
  par(mar=c(3,3,1,6),xpd=T)
  plot(e222,border="white")  
  image(fire_loc,add=T, breaks=c(0,0.99,366.5,10000),col=c("white","red","white"))
  plot(e222,add=T)
  plot(aus_reg2,border="grey",add=T)
  plot(mdb_reg2,border="grey",lwd=3,add=T)
#   text(1430000,-1000000,"study region",cex=1.2)
  text(149.1, -35.3,"ACT",font=2,cex=1.2)
  text(148.4, -36,"Kosciuszko",font=2,cex=0.9)#,srt=90)
  text(147.5, -36.4,"VIC",font=2,cex=1.2)
  text(148, -35.5,"NSW",font=2,cex=1.2)
  image.plot(zlim=c(0,1),breaks=c(0,0.5,1),lab.breaks=c("","burned \narea",""),legend.only=T,horizontal=F,col=c("red","red"),legend.shrink=0.03,legend.width=1.3,legend.args=list(text="",line=0),lwd.poly=0.1)
legend(149,-36.1,c("State","MDB"),col=c("grey","grey"),lwd=c(1,3),bty="n",title="Boundary",cex=0.9)
 dev.off()

@@

## trend of evi (MODIS product) from GA and ABARES

## tree cover data
  mod44_tc <- brick("Data\\Satellite\\MOD44\\B\\mod44_tc_aus.tif")
  tm <- seq(as.Date("2000-03-06"), as.Date("2016-12-06"), "year")
  mod44_tc <- setZ(mod44_tc, tm, "year")

## tree cover in sub-region
  mod_tc_r11 <- crop(mod44_tc,e1)
  mod_tc_r111 <- reproject(mod_tc_r11,CRS=CRS("+proj=longlat +lat_1=-18 +lat_2=-36 +lat_0=-27 +lon_0=132 +x_0=0 +y_0=0 +ellps=WGS84"))
  mod_tc_r1 <- crop(mod_tc_r111,e111)

  mod_tc_r22 <- crop(mod44_tc,e2)
  mod_tc_r222 <- reproject(mod_tc_r22,CRS=CRS("+proj=longlat +lat_1=-18 +lat_2=-36 +lat_0=-27 +lon_0=132 +x_0=0 +y_0=0 +ellps=WGS84"))
 mod_tc_r2 <- crop(mod_tc_r222,e222)

 remove(list=c("mod44_tc","mod_tc_r11","mod_tc_r111","mod_tc_r22","mod_tc_r222"))

## apply t-test to tree cover data
  fun <- function(x,m) {
    var_x <- sd(x)
    mu_1 <- mean(x[1:3])
    if (m==1) {y <- x[6:11]} else {y <- x[4:11]}   
    mu_2 <- mean(y)
    z <- (mu_2-mu_1)/var_x
    if (!is.na(z) & abs(z)>1.645) {return(y[1]-x[3])} else {return(NA)}
  }

  mod_tc_r1.z <- calc(mod_tc_r1,function(x) fun(x,1))
  mod_tc_r2.z <- calc(mod_tc_r2,function(x) fun(x,2))

  remove(list=c("mod_tc_r1","mod_tc_r2"))

## create plot
  plotclr <- brewer.pal(6,"PiYG")
#"#C51B7D" "#E9A3C9" "#FDE0EF" "#E6F5D0" "#A1D76A" "#4D9221"

  setEPS()
  postscript("tc_qld.eps")
  par(mar=c(3,3,1,7))
  plot(e111)
  image(mod_tc_r1.z,col=plotclr,breaks=c(-70,-50,-20,0,20,50,70),legend=F,add=T)
  plot(e111,add=T)
  plot(mdb_reg2,border="grey",lwd=3,add=T)
  image.plot(zlim=c(-3,3),breaks=c(-3,-2,-1,-0.2,0.2,1,2,3),lab.breaks=c(-70,-50,-20,"no tree","0 or",20,50,70),legend.only=T,horizontal=F,col=c("#C51B7D","#E9A3C9","#FDE0EF","white","#E6F5D0","#A1D76A","#4D9221"),legend.shrink=0.6,legend.width=0.7,legend.args=list(text=expression(paste(Delta,"tree (% ground)",sep="")),line=0.5,cex=0.9,font=2),axis.args=list(cex.axis=0.8),graphics.reset=F)
  dev.off()

  setEPS()
  postscript("tc_nswvic.eps")
  par(mar=c(3,3,1,7))
  plot(e222)
  image(mod_tc_r2.z,col=plotclr,breaks=c(-70,-50,-20,0,20,50,70),legend=F,add=T)
  plot(e222,add=T)
  plot(aus_reg2,border="grey",add=T)
  plot(mdb_reg2,border="grey",lwd=3,add=T)
#   image.plot(zlim=c(-3,3),breaks=c(-3,-2,-1,-0.2,0.2,1,2,3),lab.breaks=c(-70,-50,-20,"no tree","0 or",20,50,70),legend.only=T,horizontal=F,col=c("#C51B7D","#E9A3C9","#FDE0EF","white","#E6F5D0","#A1D76A","#4D9221"),legend.shrink=0.6,legend.width=0.7,legend.args=list(text=expression(paste(Delta,"tree (% ground)",sep="")),line=0.5,cex=0.7,font=2),axis.args=list(cex.axis=0.8))
# legend(149,-36.5,c("State","MDB"),col=c("grey","grey"),lwd=c(1,3),bty="n",title="Boundary",cex=0.9)
  dev.off()

setEPS()
postscript("map.eps")
par(mar=c(1.5,1.5,1.5,1.5))
plot(mdb_reg2,border="grey",lwd=3)
plot(aus_reg2,border="grey",add=T)
legend(150.2,-36,c("State","MDB"),col=c("grey","grey"),lwd=c(1,3),bty="n",title="Boundary",cex=0.9)
#invisible(text(getSpPPolygonsLabptSlots(aus_reg2), labels=as.character(aus_reg2$NAME), cex=0.7,adj=0.5))
box(lty=1)
plot(e111,col="red",add=T)
plot(e222,col="red",add=T)
text(147.2,-26.8,"QLD",cex=1.5)
text(148,-34.7,"NSW/VIC",cex=1.5)
dev.off()

@@

Specifically, this study focused on the effect of 2003 - 2004 land clearings in the QLD region and the effect of 2003 bushfires in the NSW/VIC region. These events are expected to cause a step change in the local rainfall. The actual tree cover change at the pixel level during this time was derived from the 11-year MODIS data (discussed below). The difference of tree cover before and after the land disturbance was tested using a Student's t-test. As the length of the tree cover data is shorter than the length of the rainfall data, earlier land clearings in the QLD region cannot be identified spatially, hence they are excluded from the analysis.

@@

# setwd("R:\\AVSproject\\WrittenWork\\PhD_thesis\\chap3")
# setwd("B:/My Study/PhD/AVSproject_laptop/WrittenWork/PhD_thesis/chap3")
#setwd("B:/My Study/PhD/AVSproject_PhD/WrittenWork/Paper/Paper 1 - chapter 3")

library(reshape)
library(chron)
library(fields)
library(rgdal)
library(raster)
library(maptools)
library(graphics)
library(zoo)
library(RColorBrewer) # creates nice color schemes
# library(classInt)     # finds class intervals for continuous variables

library(mgcv)
library(ncdf)
library(nnet)

@@

## predicting model: land cover is category variable
rain.trend <- function(x,y,m,var.data) {

>>>>

\begin{abstract}
Analysis of observational data to identify relationships between rainfall and land cover change are scarce due to multiple environmental factors that cannot be controlled. In this study we investigated the relationship using statistical methods on data from best available sources at two sites in Australia. Gridded data of rainfall and tree cover were used as spatially corresponding local conditions. Large scale effects were represented by climatic indicators, such as SOI and IOD. Regression analysis and step trend tests were used to assess the effect of abrupt land surface intervention. At a Queensland site, dramatic tree cover change between 2002 - 2005 did not result in strong statistically significant precipitation changes. On the other hand, results from step trend test on 1998 - 2008 rainfall data in a bushfire affected NSW/VIC region has a better match with the tree cover change map. This indicates the method works better when a abrupt change in the data can be clearly identified. The results from the step trend test implied a positive relationship between the tree cover and the rainfall at 0.1 significance level. 
\end{abstract}

@@

Land use and land cover changes can lead to changes in the local climate. Empirical and modelling studies have found cloud types and rainfall are correlated to large scale vegetation cover changes, such as deforestation in the Amazon and in the Sahel \citep{Chagnon2005,Pinto2009, Wang2009, Mei2010, kucharski_further_2013,pitman_scale_2016} and afforestation in south Israel \citep{Otterman1990, Ben-Gai1998}. Using airborne measurement in Western Australia, \citet{Junkermann2009} showed a significantly higher level of aerosols over an agricultural area compared to an adjacent natural vegetation. They suggested that a modification of aerosol concentrations due to deforestation could have contributed to a reduction of local rainfall, as more, but smaller rain droplets were observed. \citet{Nair2011} reported from the Bunny Fence Experiment in Western Australia that local land use change altered the synoptic west coast trough dynamics and surface roughness, and this resulted in an observed rainfall decrease. Maximum tempratures were also found to be sesnitive to land cover change in eastern Australia \citep{McAlpine2007}.

Overall the number of empirical studies analyzing changes to rainfall due to land cover change from observational data is limited. Most of the studies mentioned previously were either model simulations, or comparisons of modelled data with observations. This is because there are some fundamental experimental difficulties in both space (where does evaporated water reappear as rainfall?) and in time (how much time does it take for land cover change effects to appear or disappear?). In addition, in many areas across the globe rainfall variability is related to complex set of interactions, of which landuse change might only be a minor component.

Locally, there are two main sources that generate rainfall: moisture from advective atmospheric transport; and local evapotranspiration \citep{Eltahir1996,Bosilovich2006,Dirmeyer2009,Gimeno2010}. The local evapotranspiration component is the component considered to be affected by landuse change \citep{Eltahir1996}. According to \citet{Trenberth1999}, the contribution of advective moisture partially depends on the availability of external moisture and atmospheric transport. On the longer time scale, such as monthly and annually, large scale atmospheric dynamics are affected by large scale climate drivers. For example, many studies have reported significant relationships between rainfall in large parts of Australia and the El Ni\~{n}o-Southern Oscillation (ENSO) \citep{Verdon2004,Risbey2009,Speer2011}. In contrast, local ET is determined by local land surface characteristics, which influence local scale atmospheric dynamics and hence the amount of rainfall, including contribution from both main sources. Therefore land surface is suggested to play an important role in local rainfall.

%%Conventionally, large scale climate drivers are used as predictors for seasonal rainfall forecasts. According to the Australian Bureau of Meteorology (BoM), probablistic seasonal outlooks are developed based on Australian rainfall and temperature as well as sea surface temperature records from the tropical Pacific and Indian Oceans \citep{BoM2012c}. 
%% comment Expand this section with section from methods (section 4.1)?
%%Different parts of the Australian continent can be more or less influenced by different climatic drivers \citep{Mla2008}. Using both SOI and PDO in a prediction model, \citet{Kamruzzaman2011} reported that PDO was rarely significant for rainfall stations in the MDB and on the southeast coast of Australia, while SOI was at least significant at the 1\% level. In addition to SOI, \citet{Speer2011} found that the observed rainfall decrease in the southeast of NSW was linked to an increasingly positive SAM in 1976 - 1998. In southwest WA, the influence of any oceanic indices is small \citep{Smith2012}.

Although climate drivers demonstrate some capability to predict Australian rainfall, there is still a large amount of unexplained variance. \citet{Westra2010} pointed out that models based on global sea surface temperature anomalies can only predict up to 14.7\% of annual precipitation variance. Some of the remaining variance could be due to land surface processes as suggested in studies predicting local rainfall \citep[e.g.][]{Ma2011,Zeng2012, pitman_scale_2016, saha_investigating_2016}. However, they are mostly based on modelling experiments and little evidence was reported from observations. However, \citet{Pitman2004} found a good match between observations and simulated rainfall changes in southwest Western Australia ,forced by land cover change. \citet{Timbal2006} were able to reproduce the rainfall decline in south west Australia by including land cover influence. INaddititon, local land use change might not be a primary, but is likely to be a secondary cause of rainfall change \citep{Nicholls2006}.

Therefore, the aim of this study is to expand the available methodology to use empirical evidence at regional scales to investigate the cause and effect relationship between land cover change and local rainfall. More specifically, we hypothesize that a step change on the land cover on the surface will cause a step change in the rainfall. To demonstrate the approach and this effect we applied the methodlogy to study the changes in rainfall at two locations in Queensland and NSW/Victoria where there are possible step changes in land cover change due to land clearing and bushfires. The methodology is mainly based on statistical approaches to identify changes in rainfall, which were subsequently associated with land cover change through spatial comparison.

In this paper, after this section (the introduction), section 2 covers the case study areas and the observed land use change. Section 3 describes the data used in the study in more detail. Section 4 details the statistical methods and the underlying assumptions related to the modelling approach, Section 5 gives the results, which are further discussed in section 6 and finally section 7 offers the conclusions.

@@

We use two areas in Australia to demonstrate the approach. In Australia, significant tree cover change has mainly occurred in the north east of the continent and on the southeast coast, as well as in the southwest of Western Australia. According to the National Dynamic Land Cover Dataset (DLCD) \citep{Lymburner2010}, most of these areas have experienced decreasing EVI between 2000 - 2008. As an index for vegetation greenness, the decreasing values indicate lower biomass over time in the tree cover regions. The possible EVI reduction might be due to land clearing, bushfires or drought.

@@

Two regions were selected where significant tree cover change was reported. The first region is located in south central Queensland to the north of the Murray Darling Basin (MDB) (site 1 in Figure~\ref{fig:sel_reg}). High rates of land clearing have been reported in this region during the early 2000s \citep{SLATS2004}. The second study region is located at the border of New South Wales and Victoria, and includes the Snowy Mountain ranges (site 2 in Figure~\ref{fig:sel_reg}). Severe bushfires occurred in this area and the surroundings in early 2003 (see Figure~\ref{fig:bushfire}). The 2003 bushfires were the largest and the worst in this area for the last 60 years \citep{Fire2011}. Two thirds of Kosciuszko national park was heavily burned and regrowth was reported to be slow due to drought and cold conditions \citep{ABC2003} and the type of species in this region. However, in the longer term, after an early high transpiration period a recovery of pre-fire evapotranspiration would be expected \citep{Kuzcera1987}. As a result , significant tree cover loss has happened in both study areas in the last decade, either permanently or temporarily.  
%\footnote{Bushfire history, \url{http://www.dse.vic.gov.au/fire-and-other-emergencies/major-bushfires-in-victoria}, viewed on 12/12/2011}
%\footnote{Kosciuszko slow to recover from bushfires, \url{http://www.abc.net.au/news/2003-10-30/kosciuszko-slow-to-recover-from-bushfires/1501722}, viewed on 06/06/2012}

@@

<<packages,eval=T, echo = F, results = hide>>=

#library(chron)
library(fields)
library(graphics)
library(maptools)
library(plotKML)
library(raster)
library(reshape)
library(rgeos)
library(rgdal)
library(RColorBrewer)
#library(zoo)
library(tidyverse)
require(lubridate)
# library(classInt)     # finds class intervals for continuous variables
library(deseasonalize)
library(mgcv)
library(ncdf4)
library(nnet)
@

<<mapping,eval=T, echo = F, results = hide>>=

## define the boundary of sub-regions
   e1 <- extent(c(1324304.091,1723901.773,-222556.668,80451.265))
   e2 <- extent(c(1346395.141,1534425.527,-1169402.246,-607644.171))
  # 
   e11 <- extent(c(145.425, 149.025, -28.475, -25.175))
   e22 <- extent(c(146.675, 149.325, -37.075, -34.675))

  e111 <- extent(c(145.825, 148.625, -28.075, -25.575)) 
  e222 <- extent(c(147.075, 148.925, -36.675, -35.075))

## the selected regions for mapping
root_dir <- "c:/users/rver4657/owncloud/chun/chun/data"
aus_reg <- readShapePoly(paste(root_dir,
                               "Geo\\aus_state2011\\STE11aAust.shp",sep="/"),
                         proj4string=CRS("+proj=lcc +lat_1=-18 +lat_2=-36 +lat_0=-27 +lon_0=132 +x_0=0 +y_0=0"))
aus_reg2 <- gSimplify(aus_reg,0.01,topologyPreserve=T)

mdb_reg <- readShapePoly(paste(root_dir,
                               "Geo\\MDB\\mdb_boundary.shp", sep="/"),
                               proj4string=CRS("+proj=lcc +lat_1=-18 +lat_2=-36 +lat_0=-27 +lon_0=132 +x_0=0 +y_0=0"))
mdb_reg2 <- gSimplify(mdb_reg,0.01,topologyPreserve=TRUE)

#remove(list=c("aus_reg","mdb_reg"))

@@

fire_0301 <- raster(paste(root_dir,
                          "Other\\bushfires\\Burnt_Area\\tmpA2003001.burndate.tif",sep="/"))
fire_0301nsw <- crop(fire_0301,e2)
fire_loc <- crop(reproject(fire_0301nsw,CRS=CRS("+proj=longlat +lat_1=-18 +lat_2=-36 +lat_0=-27 +lon_0=132 +x_0=0 +y_0=0 +ellps=WGS84")),e222)
remove(list=c("fire_0301","fire_0301nsw"))

  jpeg("../figures/bushfire_nswvic.jpg", width=960, height=960)
  par(mar=c(3,3,1,6),xpd=T)
  plot(e222,border="white")  
  image(fire_loc,add=T, breaks=c(0,0.99,366.5,10000),col=c("white","red","white"))
  plot(e222,add=T)
  plot(aus_reg2,border="grey",add=T)
  plot(mdb_reg2,border="grey",lwd=3,add=T)
#   text(1430000,-1000000,"study region",cex=1.2)
  text(149.1, -35.3,"ACT",font=2,cex=1.2)
  text(148.4, -36,"Kosciuszko",font=2,cex=0.9)#,srt=90)
  text(147.5, -36.4,"VIC",font=2,cex=1.2)
  text(148, -35.5,"NSW",font=2,cex=1.2)
  image.plot(zlim=c(0,1),breaks=c(0,0.5,1),lab.breaks=c("","burned \narea",""),legend.only=T,horizontal=F,col=c("red","red"),legend.shrink=0.03,legend.width=1.3,legend.args=list(text="",line=0),lwd.poly=0.1)
legend(149,-36.1,c("State","MDB"),col=c("grey","grey"),lwd=c(1,3),bty="n",title="Boundary",cex=0.9)
 dev.off()

@@

rootdir <-"C:/users/rver4657/owncloud/chun2/data/satellite/MOD44/B"

# read in the file list from the directory
x <- dir(path=rootdir,pattern="Percent_Tree_Cover.tif")
# Identify the different states in the images
# NSW
NSW <- x[grep("NSW",x)]
# Qld
Qld <- x[grep("Qld",x)]

img <- list()
# raster and brick
for(i in 1:length(NSW)) {
  img[[i]] <- raster(paste(rootdir,NSW[i],sep="/"))
}

imgQ <- list()
# raster and brick
for(i in 1:length(Qld)) {
  imgQ[[i]] <- raster(paste(rootdir,Qld[i],sep="/"))
}

NSW_b <- do.call(brick,img)
Qld_b <- do.call(brick, imgQ)

tm <- seq(as.Date("2000-03-06"), as.Date("2016-03-06"), "year")
mod_tc_r22 <- setZ(NSW_b, tm, "year")

mod_tc_r11 <- setZ(Qld_b, tm, "year")

## apply t-test to tree cover data
fun <- function(x,m) {
  var_x <- sd(x,na.rm=T)
  mu_1 <- mean(x[1:3],na.rm=T)
  # m = 1 Queensland, 
  if (m==1) {y <- x[4:16]} else {
    # NSW whole series m = 2
    if (m ==2) {y <- x[6:16]} else {
      # build in other possibilities for length series
      y <- x[6:(16 - m)]
    }  
  mu_2 <- mean(y,na.rm=T)
  z <- (mu_2-mu_1)/var_x
  if (!is.na(z) & abs(z)>1.645) {return(mu_2-mu_1)} else {return(NA)}
  }
}

# Qld
mod_tc_r1.z <- calc(mod_tc_r11,function(x) fun(x,1))
# NSW
mod_tc_r2.z <- calc(mod_tc_r22,function(x) fun(x,2))
# NSW only 5 years after
mod_tc_r2_5y.z <- calc(mod_tc_r22,function(x) fun(x,7))
# NSW 7 years after
mod_tc_r2_7y.z <- calc(mod_tc_r22,function(x) fun(x,5))
# NSW 9 years after
mod_tc_r2_9y.z <- calc(mod_tc_r22,function(x) fun(x,3))

require(RColorBrewer)
plotclr <- brewer.pal(8,"PiYG")

jpeg("../figures/tc_nswvic_5y.jpg", width=960, height=960)
par(mar=c(3,3,1,7))
plot(e222)
image(mod_tc_r2_5y.z,col=plotclr,breaks=c(-70,-20,-10,-5,0,5,10,20,70),legend=F,add=T)
#plot(e111,add=T)
plot(aus_reg2,border="grey",lwd=3,add=T)
image.plot(zlim=c(-4,4),breaks=c(-4,-3,-2,-1,-0.2,0.2,1,2,3,4),
           lab.breaks=c(-70,-20,-10,-5,"no tree","0 or",5,10,20,70),
           legend.only=T,horizontal=F,
           col=c("#C51B7D","#DE77AE","#F1B6DA","#FDE0EF","white",
                 "#E6F5D0","#B8E186","#7FBC41","#4D9221"),
           legend.shrink=0.7,
           legend.width=0.7,
           legend.args=list(text=expression(paste(Delta,
                            "tree (% ground)",sep="")),
                            line=0.5,cex=0.9,font=2),
           axis.args=list(cex.axis=0.8),graphics.reset=F)
dev.off()

jpeg("../figures/tc_nswvic_7y.jpg", width=960, height=960)
par(mar=c(3,3,1,7))
plot(e222)
image(mod_tc_r2_7y.z,col=plotclr,breaks=c(-70,-20,-10,-5,0,5,10,20,70),legend=F,add=T)
#plot(e111,add=T)
plot(aus_reg2,border="grey",lwd=3,add=T)
image.plot(zlim=c(-4,4),breaks=c(-4,-3,-2,-1,-0.2,0.2,1,2,3,4),
           lab.breaks=c(-70,-20,-10,-5,"no tree","0 or",5,10,20,70),
           legend.only=T,horizontal=F,
           col=c("#C51B7D","#DE77AE","#F1B6DA","#FDE0EF","white",
                 "#E6F5D0","#B8E186","#7FBC41","#4D9221"),
           legend.shrink=0.7,
           legend.width=0.7,
           legend.args=list(text=expression(paste(Delta,
                            "tree (% ground)",sep="")),
                            line=0.5,cex=0.9,font=2),
           axis.args=list(cex.axis=0.8),graphics.reset=F)
dev.off()

jpeg("../figures/tc_nswvic_9y.jpg", width=960, height=960)
par(mar=c(3,3,1,7))
plot(e222)
image(mod_tc_r2_9y.z,col=plotclr,breaks=c(-70,-20,-10,-5,0,5,10,20,70),legend=F,add=T)
#plot(e111,add=T)
plot(aus_reg2,border="grey",lwd=3,add=T)
image.plot(zlim=c(-4,4),breaks=c(-4,-3,-2,-1,-0.2,0.2,1,2,3,4),
           lab.breaks=c(-70,-20,-10,-5,"no tree","0 or",5,10,20,70),
           legend.only=T,horizontal=F,
           col=c("#C51B7D","#DE77AE","#F1B6DA","#FDE0EF","white",
                 "#E6F5D0","#B8E186","#7FBC41","#4D9221"),
           legend.shrink=0.7,
           legend.width=0.7,
           legend.args=list(text=expression(paste(Delta,
                            "tree (% ground)",sep="")),
                            line=0.5,cex=0.9,font=2),
           axis.args=list(cex.axis=0.8),graphics.reset=F)
dev.off()

jpeg("../figures/tc_nswvic.jpg", width=960, height=960)
par(mar=c(3,3,1,7))
plot(e222)
image(mod_tc_r2.z,col=plotclr,breaks=c(-70,-20,-10,-5,0,5,10,20,70),legend=F,add=T)
#plot(e111,add=T)
plot(aus_reg2,border="grey",lwd=3,add=T)
image.plot(zlim=c(-4,4),breaks=c(-4,-3,-2,-1,-0.2,0.2,1,2,3,4),
           lab.breaks=c(-70,-20,-10,-5,"no tree","0 or",5,10,20,70),
           legend.only=T,horizontal=F,
           col=c("#C51B7D","#DE77AE","#F1B6DA","#FDE0EF","white",
                 "#E6F5D0","#B8E186","#7FBC41","#4D9221"),
           legend.shrink=0.7,
           legend.width=0.7,
           legend.args=list(text=expression(paste(Delta,
                            "tree (% ground)",sep="")),
                            line=0.5,cex=0.9,font=2),
           axis.args=list(cex.axis=0.8),graphics.reset=F)
dev.off()

jpeg("../figures/tc_qld.jpg", width=960, height=960)
par(mar=c(3,3,1,7))
plot(e111)
image(mod_tc_r1.z,col=plotclr,breaks=c(-70,-20,-10,-5,0,5,10,20,70),legend=F,add=T)
#plot(e111,add=T)
plot(aus_reg2,border="grey",lwd=3,add=T)
image.plot(zlim=c(-4,4),breaks=c(-4,-3,-2,-1,-0.2,0.2,1,2,3,4),
           lab.breaks=c(-70,-20,-10,-5,"no tree","0 or",5,10,20,70),
           legend.only=T,horizontal=F,
           col=c("#C51B7D","#DE77AE","#F1B6DA","#FDE0EF","white",
                 "#E6F5D0","#B8E186","#7FBC41","#4D9221"),
           legend.shrink=0.7,
           legend.width=0.7,
           legend.args=list(text=expression(paste(Delta,
                            "tree (% ground)",sep="")),
                            line=0.5,cex=0.9,font=2),
           axis.args=list(cex.axis=0.8),graphics.reset=F)
dev.off()

jpeg("../figures/map.jpg", width=960, height=960)
par(mar=c(1.5,1.5,1.5,1.5))
plot(mdb_reg2,border="grey",lwd=3)
plot(aus_reg2,border="grey",add=T)
legend(150.2,-36,c("State","MDB"),col=c("grey","grey"),lwd=c(1,3),bty="n",title="Boundary",cex=0.9)
#invisible(text(getSpPPolygonsLabptSlots(aus_reg2), labels=as.character(aus_reg2$NAME), cex=0.7,adj=0.5))
box(lty=1)
plot(e111,col="red",add=T)
plot(e222,col="red",add=T)
text(147.2,-26.8,"QLD",cex=1.5)
text(148,-34.7,"NSW/VIC",cex=1.5)
dev.off()

@@

The hypothesis in this study therefore is that the effect of 2003 - 2004 land clearings in the QLD region and the 2003 bushfires in the NSW/VIC region cause a step change in the local rainfall. The actual tree cover change at the pixel level during this time was derived from the 15-year MODIS data (discussed below). The difference of tree cover before and after the land disturbance was tested using a Student's t-test. As the length of the tree cover data is shorter than available rainfall data, earlier land clearings in the QLD region cannot be identified spatially, hence they are excluded from the analysis.

@@

@@

## predicting model: land cover is category variable
rain.trend <- function(x,y,var.data) {

  var.data$rain <- as.vector(unlist(y[x]))[1:nrow(var.data)]+0.0001
